---
title: "Weight Lifting Exercises Dataset"
output: html_document
---


```{r test,echo=FALSE, cache=TRUE}
### load the data in ###
input.train.dat <- read.csv("pml-training.csv", header=T)
input.valid.dat <- read.csv("pml-testing.csv", header=T)


```


# Data Overview

This human activity recognition research collected 6 participants to record five ways in performance of barbell lifts that measured from meters on the belt, forearm, arm, and dumbbell. The 5 ways of barbell lifts activities (column name: "classe", denoted as A-E) were (<http://groupware.les.inf.puc-rio.br/har>):

    1. exactly according to the specification (Class A),
    2. throwing the elbows to the front (Class B), 
    3. lifting the dumbbell only halfway (Class C), 
    4. lowering the dumbbell only halfway (Class D), and 
    5. throwing the hips to the front (Class E).

The goal for this study is to predict what type of activities from accelerometers data on the belt, forearm, arm, and dumbell. 

### Preparing data

We split 'pml-training' data with 60% samples as a training data, the rest 40% sample as a testing data to test our model. The validation dataset will be data 'pml-testing'.


```{r}
library(caret)
### get data from accelerometers on the belt, forearm, arm, and dumbell
use.column <- c(grep("user_name", names(input.train.dat)), grep("^accel_", names(input.train.dat)),grep("^total_accel_", names(input.train.dat)), grep("classe", names(input.train.dat)))


use.train.dat <- input.train.dat[ , use.column]

#### create training and testing dataset in training data ###
set.seed(12345)
inTrain <- createDataPartition(y = use.train.dat$classe, p = 0.6, list = FALSE)

train.dat <- use.train.dat[inTrain, ]
test.dat <- use.train.dat[-inTrain, ] ### independent dataset for accuracy 
```


### PCA for preprocessing data

Principle component analysis (PCA) was applied for preprocessing the predictors. Based on our training data, we need 10 principle components to capture 95% variability.

```{r preproc}
### preprocessing with pca by creating new variables that captured 95% variability###
col.usename.classe <- seq(1, dim(train.dat)[2])[names(train.dat) %in% c("user_name", "classe")]

pre.proc <- preProcess(train.dat[,-col.usename.classe], method = "pca",thresh = 0.95) 
pre.proc ### See how many variables needed 
```



### Build Random forest model with cross validation to test our model prediction

We built the model via `random forest` procedure with 5-fold cross validation. 
```{r rfmodel}
train.PC.var <- predict(pre.proc, train.dat[,-col.usename.classe])
train.PC.var$classe <- train.dat$classe
modelFit.pc.train <- train(classe ~ . , method = "rf", data = train.PC.var, trControl=trainControl(method = "cv", number = 5))

```

###  see model accuracy from testing data
we applied our prediction model to the testing data for accuracy estimator.

```{r indval}
test.dat.PC.var <- predict(pre.proc, test.dat[,-col.usename.classe])
test.dat.PC.var$classe <- test.dat$classe
pred.result.pc.test <- predict(modelFit.pc.train, test.dat.PC.var)

confusionMatrix(test.dat$classe, pred.result.pc.test)
```

### Apply the developed model to validation dataset

The predicted classe from our model is:

```{r valid, echo=FALSE}

use.valid.dat <- input.valid.dat[, names(input.valid.dat) %in% names(test.dat)] 

use.valid.dat.afterPC <- predict(pre.proc, use.valid.dat[ , -c(1)])
predict(modelFit.pc.train, use.valid.dat.afterPC)

```
